---
title: "Returns for Stocks and Bonds under each Presidential Party"
author: "Michael Boerman"
output: html_notebook
---

This is a simple code for showing how stocks and bonds do when the presidential party is Democratic or Republican. This is less about analysis (arrow of causation, anyone?) and more of a *very* simple showcase of my tidyverse toolbox!

This code is 3 chunks:

1)  Libraries

2)  Read data

3)  Plot!

Chunks 2 and 3 are each one call. You can execute by using `ctrl + shift + enter` after clicking inside the plot. Run the entire file by using `ctrl + alt + r`. You will most likely need to install packages for the first chunk.

Please recall that correlation != causation :). I am not registered or associated with any political party.

### 1) Libraries

```{r, message = FALSE, warning = FALSE}

# The usuals!
library(readxl)     # for reading the excel file
library(tidyverse)  # all inclusive
library(here)       # for relative file paths
library(scales)     # for easy conversion of decimal to % in plots
library(ggtext)     # for spicing up the text with color.
library(png)        # to read in the Github logo
library(grid)       # to render  the Github logo
library(extrafont)  # for special fonts

# --- font setup --- #
# run this *once*; it takes awhile! then comment-out.
# font_import()
loadfonts(device="win") 

# prevent intermediate messages/outputs from showing up
# after each chunk:
knitr::opts_chunk$set(message = FALSE)

```

### 2) Read Data

The data was given to me by [Jacob Stephens](https://jacobcstephens.com/about/). He collected from (?). The spreadsheet can be [downloaded from his sharepoint.](https://gordonedu-my.sharepoint.com/:x:/g/personal/jacob_stephens_gordon_edu/EVx2b0Dwb3JJtNbFIklmS-gB_ZZAI9ei7jH0Bdvb6yWAEQ?e=eWGy8Y) *(Unfortunately, it can't be read from a url due to the way sharepoint hosts files)*.

```{r}
# Load the data!

# First, read from a local csv.
data <-
  read_xlsx(here("data_input/president-bond-stock-change-correlation.xlsx"),
    skip = 4
  ) %>%

  # Next, rename the columns.
  # excel saves spaces stupidly, so need this weird dot/space
  #  (don't ask how long i spent figuring this out!)
  rename(
    year   = ` Year`,
    party  = Party,
    bonds  = `100% Bonds `,
    stocks = `S&P 500 Index `,
  ) %>% 
  
  # Next, insert a color for use in plotting later. Ignore case and endings (ie: read "democrat" the same as "Democratic".)
  # I'll use the official colors of Repub and Dem parties
  mutate(political_colors = case_when(
    str_detect(party, "emocrat") ~ "#0015BC",
    str_detect(party, "epublic") ~ "#DE0100"
  )) %>% 
  
  # Next, some data manipulation. Select just a few columns...
  dplyr::select(party, year, bonds, stocks, political_colors) %>% 
  
  # and turn these into "long" data instead of wide.
  # this is needed for proper plotting!
  pivot_longer(
    cols = c(bonds, stocks), 
    names_to = "asset", 
    values_to = "return"
  )

# write the intermediate output.
write_csv(data, here("results/csv/cleaned_data.csv"))
```

A quick aside: I'd like to use the Github logo in the chart, but am not a fontawesome pro member. Thus, I downloaded the png and will import it:

```{r}
github_logo <- readPNG(here("data_input/github_logo.png")) %>%
  grid::rasterGrob()
                    
```

### 3) Plots!

I had been using pmap with several inputs, but using `pivot_longer` above eliminated this need.

```{r, fig.show="hide",  results="hide", fig.keep = "none"}

    political_colors <- unique(data$political_colors)
    ## ---  parameter set up: --- ##

    # clean up the ggplot code later on by naming them now
    plot_title <- "Which party throws the best market party?"
    
    plot_subtitle <- 
    "Annual market return under <strong><span style='color:#DE0100'>Repbulican presidents</span></strong> varies widely, \n while the returns under <strong><span style ='color:#0015BC'>Democratic presidents</span></strong> have fewer negative years."
   

   data_source <- "S&P 500 & Bond returns (mix of U.S. Treasury bond durations and TIPS)"

   plot_caption <- paste0(
     "Data are annual ", data_source, ", from ",
     min(data$year), " to ", max(data$year), "."
   )
  
    
    ## --- Means --- ##

    # # grab the means for each party for each asset class
    # party_means <- data %>%
    #   select(party, return, asset) %>%
    #   group_by(party) %>%
    #   summarize(
    #     stock_return = where(asset == "stock" = mean(return),
    #     bond_return  = mean(return),
    #     .groups = "drop"
    #   ) %>%
    #   select(-party) %>%
    #   identity()
    
    ## --- Plots! --- ##

    ### Plot #1: Boxplot
    data %>% 
      ggplot(aes(x = asset, y = return, color = political_colors)) +
      scale_color_manual(values = political_colors) +
      geom_violin(
        alpha = 0.5,
        size  = 1,
        # dodge.width = 0.75
      ) +
      ggbeeswarm::geom_quasirandom(
        # shape       = 1,
        size        = 3,
        width       = 0.15,
        dodge.width = 0.85,
        alpha       = 0.75,
        show.legend = F
      ) +
      scale_y_continuous(
        labels = percent_format(accuracy = 2),
      ) +    
      scale_x_discrete(labels = c("Bonds", "Stocks")) +
      theme_minimal() +
      labs(
        title    = plot_title,
        subtitle = plot_subtitle,
        caption  = plot_caption
      ) +
      theme(
        legend.position = "none",
        axis.title      = element_blank(),
        plot.subtitle   = element_markdown(size = 12),
        plot.title      = element_text(size = 18),
        text = element_text(
          family = "Times New Roman",
          size = 16
        ),
        plot.caption = element_text(size = 10)
      ) +
      
      # placing the Github logo, courtesy of this great guide:
      # https://www.markhw.com/blog/logos
      annotation_custom(
        github_logo,
        xmin = 0.1, xmax = 1.1,    # manual location represents ever
        ymin = -.3, ymax = -.205   #-thing I hate about life.
      ) +
      annotate(
        "text", 
        x = 1.10, y = -.25, 
        size = 10,
        label = "/michaelboerman",
        family = "Courier New"
      ) +
      geom_hline(yintercept = 0) 
      
    ggsave(filename = here("results/plots/asset_returns.png"), 
             width = 10, height = 6)


```
