---
title: "Vector Autoregressive Analysis"
author: "Michael Boerman"
date: "4/8/2021"
output: html_document
---

In this second analysis, I'd like to take a look at the time-delayed impact of presidential political party on stock and bond returns.

To do this, I'll use vector autoregressions. While simple regressions assume uni-directional influence, VARs treat all variables endogenously. This allows us to test Granger causality, too, which specifies if one variable is useful in forecasting another. For a fuller discussion, see the legendary[Rob Hyndman's slides](https://robjhyndman.com/eindhoven/3-4-Advanced.pdf).

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(stringr)

library(fpp)
library(vars)
library(here)
```

Import the cleaned data from the first analysis, `political_stocks_plots.Rmd`:

```{r}
data <- read_csv(here("results/csv/cleaned_data.csv")) %>% 
  pivot_wider(
    id_cols = c(year, party), 
    names_from = asset, 
    values_from = return
)
```

Great! Now, some rudimental var:

```{r}
data_ts <- ts(data = cbind(data$bonds, data$stocks, data$party), start = min(data$year), end = max(data$year), frequency = 1, names = c("stock_return", "bond_return", "party"))

ar(data_ts, order=3)

VARselect(data_ts, lag.max = 4, type = "const")

var <- VAR(data_ts, p=5, type="both")
summary(var)

plot(forecast(var), xlab = "Year")

```

Granger test:

```{r}
grangertest(data$year, data$stocks, order = 10)
```
